---
title: "EPI 590R Final Project"
author: "Kelly Longhini"
format:
  html:
    embed-resources: true
execute: 
  echo: false
---

```{r}
#| warning: false
#| message: false

# Install and call packages

install.packages("tidyverse")
install.packages("dplry")
install.packages("rmarkdown")
install.packages("gtsummary")
install.packages("here")
install.packages("broom")
install.packages("broom.helpers")
install.packages("ggplot2")

library(tidyverse)
library(dplyr)
library(rmarkdown)
library(gtsummary)
library(here)
library(broom)
library(broom.helpers)
library(ggplot2)
```

## Description of the Data

The data comes from.....

```{r}
# Load in the catheterization R data set using the here package
load(here::here("data", "raw", "cath.rda")) 

cath <- cath %>%
  mutate(sex_cat = factor(cath$sex, labels = c("Male", "Female")),
         sigdz_cat = factor(cath$sigdz, labels = c("No", "Yes")),
         tvdlm_cat = factor(cath$tvdlm, labels = c("No", "Yes"))
  )

```

## Descriptive Statistics

```{r}
#| label: tbl-one
#| tbl-cap: "Table of descriptive statistics"

table1<- tbl_summary(
  cath,
  by = sex_cat,
  include = c(sex_cat, age, cad_dur, choleste, sigdz_cat, tvdlm_cat),
  label = list(
    age ~ "Age (years)",
    cad_dur ~ "Duration of chest pain symptoms (days)",
    choleste ~ "Serum cholesterol level (milligrams per deciliter)",
    sigdz_cat ~ "Significant coronary disease (at least 75% blockage of at least one of the major coronary arteries)",
    tvdlm_cat ~ "Three vessel disease/left main disease (blockage of either all three of the coronary vessels, or blockage of the left main coronary artery)"
  ),
  missing_text = "Missing") %>%
  add_overall(last = TRUE)

table1
```

@tbl-one displays descriptive statistics for the predictor and outcome variables in the data set, including age, duration of chest pain symptoms, serum cholesterol level, and coronary artery and coronary vessel blockages .

The overall median duration of chest pain symptoms is `r median(cath$cad_dur, na.rm = T)` days.

```{r}
cholesterol_median_male <- inline_text(table1, variable = "choleste", 
                         column = "stat_1")
cholesterol_median_female <- inline_text(table1, variable = "choleste", 
                         column = "stat_2")
```

The median serum cholesterol level for males (`r cholesterol_median_male` mg/dL) is lower than the median serum cholesterol level for females (`r cholesterol_median_female` mg/dL) in the study.


## Logistic Regression
```{r}
logistic_model <- glm(sigdz ~ age + sex_cat + cad_dur + choleste,
											data = cath, family = binomial())
```

```{r}
#| label: tbl-two
#| tbl-cap: "Logistic regression results"

regression <- tbl_regression(
  logistic_model,
  exponentiate = TRUE,
  label = list(
    age ~ "Age (years)",
    sex_cat ~ "Sex",
    cad_dur ~ "Duration of chest pain (days)",
    choleste ~ "Serum cholesterol level (mg/dL)"
  )
)

regression

```

@tbl-two shows the regression of significant coronary disease on a series of predictor variables, including age, sex, duration of chest pain, and serum cholesterol level. 


# Histogram of Cholesterol
```{r}
#| warning: false
#| message: false
#| label: fig-one
#| fig-cap: "Histogram of serum cholesterol level"
hist <- ggplot(cath, aes(x = choleste)) +
  geom_histogram(fill = "lightyellow", color = "black") +
  labs(title = "Histogram of serum cholesterol level",
       x = "Serum cholesterol level (mg/dL)",
       y = "Frequency")
hist

# Save the figure
ggsave(plot = hist, 
       filename = here::here("figures", "histogram.pdf"))
```

@fig-one depicts a histogram of serum cholesterol levels (mg/dL) for all participants in the study. 

# Function 






